# Tables

```{r tab_path, include = FALSE}
library(gt)
library(tibble)
library(knitr)
library(kableExtra)
# Source analysis script
source(here::here("outputs", "technical_report", "scripts", "analysis_script.R"))

```


## Table 1. Eligibility criteria for including and escluding studies in the scoping review

```{r t1, echo = FALSE, message = FALSE, warning = FALSE}
# Define PICO elements
population_include    <- "School stakeholders"
population_exclude    <- "None"

concept_include       <- "Four-day school week"
concept_exclude       <- "Other changes to the traditional school calendar (e.g., extended school year)"

context_include       <- "Primary and secondary schools in the United States"
context_exclude       <- "Pre-schools, postsecondary schools, and schools outside the United States"

design_include        <- "Any empirical design"
design_exclude        <- "Conceptual and opinion articles"

publications_include  <- "Research reports in English"
publications_exclude  <- "Reports in a language other than English and non-research reports (e.g., popular media)"

# Assemble into a data frame
criteria_df <- tibble(
  Criteria = c("Population",
               "Concept",
               "Context",
               "Study Design",
               "Publications"),
  Include  = c(population_include,
               concept_include,
               context_include,
               design_include,
               publications_include),
  Exclude  = c(population_exclude,
               concept_exclude,
               context_exclude,
               design_exclude,
               publications_exclude)
)

# Render as an HTML table
criteria_df %>%
  kable(
    format    = "html",
    align     = c("l","l","l"),
    escape    = TRUE
  ) %>%
  kable_styling(
    full_width = FALSE,
    position   = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  )

```


## Table 2. Characteristics of included sources of evidence on the four-day school week {#tbl1-study-char}

```{r t2,  echo = FALSE, message = FALSE, warning = FALSE}
# Helper function
add_table2_label <- function(df, section) {
  df %>%
    mutate(
      Section      = section,
      Characteristic = stringr::str_to_sentence(Category),
      Median       = NA_character_,
      Range        = NA_character_,
      indent       = TRUE
    ) %>%
    rename(N = N, Percent = Percentage) %>%
    select(Section, Characteristic, N, Percent, Median, Range, indent)
}

pubtype_df <- publication_summary_specific %>%
  add_table2_label("Publication Type") %>%
  mutate(
    Characteristic = case_when(
      Characteristic == "Peer-reviewed journal article" ~ "Journal",
      Characteristic == "Student dissertation or thesis" ~ "Student",
      TRUE ~ Characteristic
    )
  )

studenttype_df <- add_table2_label(student_type_summary, "Student Type")

discipline_df  <- add_table2_label(student_disc_summary, "Student Discipline")

community_df <- community_summary %>%
  add_table2_label("Community Type") %>%
  mutate(
    Characteristic = if_else(
      Characteristic == "Unknown",
      "Not reported",
      Characteristic
    )
  ) %>% 
  arrange(
    Characteristic == "Not reported"  # FALSE first, TRUE (Not reported) last
  )

school_df <- school_summary %>%
  add_table2_label("School Level") %>%
  mutate(
    Characteristic = if_else(
      Characteristic == "Unknown",
      "Not reported",
      Characteristic
    )
  ) %>% 
  arrange(
    Characteristic == "Not reported" 
  )

race_df <- race_summary %>%
  add_table2_label("Majority Student Race/Ethnicity") %>%
  mutate(
    Characteristic = case_when(
      Characteristic == "Hispanic, latino, or spanish" ~ "Hispanic, Latino, or Spanish",
      Characteristic == "American indian and/or alaska native" ~ "American Indian or Alaska Native",
      Characteristic == "Native hawaiian/other pacific islander" ~ "Native Hawaiian or Other Pacific Islander",
      TRUE ~ Characteristic
    )
  ) %>% 
  arrange(
    Characteristic == "Not reported"  
  )

fifthday_df <- fifthday_summary %>%
  add_table2_label("Fifth Day Activities") %>%
  mutate(
    Characteristic = case_when(
      Characteristic == "Nothing shut down school" ~ "Nothing (school closed)",
      Characteristic == "Teacher in service" ~ "Teacher in-service",
      TRUE ~ Characteristic
    )
  ) %>% 
  arrange(
    Characteristic == "Not reported"  
  )

# Numeric summaries get Section = "" (empty string)
years_df <- tribble(
  ~Section, ~Characteristic, ~N, ~Percent, ~Median, ~Range, ~indent,
  "", "Publication Year", NA, NA,  as.character(publication_year_summary$Median),   
  publication_year_summary$Range,  FALSE,
  "",  "Study Year", NA, NA, as.character(summary_data_years$Median),
  summary_data_years$Range, FALSE
)

# Bind rows and tidy
table2_df <- bind_rows(
  pubtype_df,
  studenttype_df,
  discipline_df,
  years_df,
  community_df,
  school_df,
  race_df,
  fifthday_df
) %>%
  mutate(
    N       = ifelse(is.na(N),       "", as.character(N)),
    Percent = as.numeric(Percent),
    Percent = ifelse(is.na(Percent),"",
                     paste0(round(Percent, 0), "%")),
    Median  = replace_na(Median,  ""),
    Range   = replace_na(Range,   "")
  )

# Build gt table
gt_tbl <- table2_df %>%
  gt(groupname_col = "Section") %>%
  cols_label(
    Characteristic = md("**Characteristic**"),
    N              = md("**N**"),
    Percent        = md("**%**"),
    Median         = md("**Median**"),
    Range          = md("**Range**")
  ) %>%
  tab_style(
    style = cell_text(indent = px(15)),
    locations = cells_body(rows = indent == TRUE, columns = "Characteristic")
  ) %>%
  # bold all the group headers (the non-empty Sections)
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups(everything())
  ) %>%
  # bold just those two numeric rows
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      rows    = Characteristic %in% c("Publication Year", "Study Year"),
      columns = "Characteristic"
    )
  ) %>%
  # hide the blank‐string group header
  tab_style(
    style = cell_text(color = "white"),
    locations = cells_row_groups(groups = "")
  ) %>%
  # remove its top border
  tab_style(
    style = cell_borders(sides = "top", color = "white", weight = px(0)),
    locations = cells_row_groups(groups = "")
  ) %>%
   # first footnote group
  tab_footnote(
    footnote = "Denominator for percentages is n = 70 student manuscripts.",
    locations = cells_row_groups(
      groups = c("Student Type", "Student Discipline")
    )
  ) %>%
  
  # second footnote group
  tab_footnote(
    footnote = "Percentages total more than 100% because categories are not mutually exclusive.",
    locations = cells_row_groups(
      groups = c("Community Type", "School Level", "Fifth Day Activities")
    )
  ) %>%
  tab_style(
    style = cell_borders(
      sides  = "bottom",
      color  = "black",
      weight = px(2)      
    ),
    locations = cells_column_labels()
  ) %>% 
  tab_options(table.font.size = "small") %>%
  cols_hide(columns = c("indent"))

gt_tbl

```



## Table 3. Types of available research evidence on the four-day school week {#tbl2-study-evidence}

```{r t3, echo = FALSE, message = FALSE, warning = FALSE}
# Define group‐header labels as simple objects (easy to update)
hdr_effectiveness   <- "Effectiveness"
hdr_acceptability   <- "Acceptability"
hdr_stakeholder     <- "Interest-Holders"
hdr_inferential     <- "Inferential Approach"
hdr_outcome         <- "Outcome Domain"
hdr_equity          <- "Equitability of Effects"

# Build each block

## The first two summary rows come from evidence_summary
eff_block <- evidence_summary[1, ] %>%
  transmute(
    Characteristic = hdr_effectiveness,
    N              = N,
    Percent        = round(Percentage, 0),
    indent         = FALSE
  )

acc_block <- evidence_summary[2, ] %>%
  transmute(
    Characteristic = hdr_acceptability,
    N              = N,
    Percent        = round(Percentage, 0),
    indent         = FALSE
  )

## Stakeholder group: header + indented subrows
stake_hdr <- tibble(
  Characteristic = hdr_stakeholder,
  N              = NA_integer_,
  Percent        = NA_real_,
  indent         = FALSE
)

stake_blk <- stakeholder_summary %>%
  transmute(
    Characteristic = str_to_sentence(Category),
    N              = N,
    Percent        = round(Percentage, 0),
    indent         = TRUE
  )

## Inferential approach
inf_hdr <- tibble(
  Characteristic = hdr_inferential,
  N              = NA_integer_,
  Percent        = NA_real_,
  indent         = FALSE
)

inf_blk <- approach_summary %>%
  transmute(
    Characteristic = str_to_sentence(Category),
    N              = N,
    Percent        = round(Percentage, 0),
    indent         = TRUE
  ) %>% 
  mutate(
    Characteristic = case_when(
      Characteristic == "Between-group (without controls)" ~ "Between-group (no statistical control)",
      Characteristic == "Between-group (with controls)" ~ "Between-group (with statistical controls)",
      TRUE ~ Characteristic
    )
  )

## Outcome domain
out_hdr <- tibble(
  Characteristic = hdr_outcome,
  N              = NA_integer_,
  Percent        = NA_real_,
  indent         = FALSE
)

out_blk <- outcome_summary %>%
  transmute(
    Characteristic = str_to_sentence(Category),
    N              = N,
    Percent        = round(Percentage, 0),
    indent         = TRUE
  ) %>% 
  mutate(
    Characteristic = case_when(
      Characteristic == "Student achievement" ~ "Achievement",
      Characteristic == "Student attendance" ~ "Attendance",
      Characteristic == "Teacher and staff recruitment and retention" ~ "Staff/Teacher retention",
      Characteristic == "Student health status or behaviors" ~ "Health",
      Characteristic == "Student attainment" ~ "Attainment",
      Characteristic == "Disciplinary incidents during school" ~ "School disciplinary incidents",
      Characteristic == "Juvenile crime" ~ "Criminal activity",
      TRUE ~ Characteristic
    )
  )

## Equitability of effects (header comes from evidence_summary[3,])
eq_hdr <- evidence_summary[3, ] %>%
  transmute(
    Characteristic = hdr_equity,
    N              = N,
    Percent        = round(Percentage, 0),
    indent         = FALSE
  )

eq_blk <- equity_summary %>%
  transmute(
    Characteristic = str_to_sentence(Category),
    N              = N,
    Percent        = round(Percentage, 0),
    indent         = TRUE
  ) %>% 
  mutate(
    Characteristic = case_when(
      Characteristic == "Student age" ~ "Age or grade level",
      Characteristic == "Community type or rurality" ~ "Community type (rurality)",
      Characteristic == "Student socioeconomic status" ~ "Socioeconomic status",
      Characteristic == "Student race or ethnicity" ~ "Race or ethnicity",
      Characteristic == "Student gender identity" ~ "Sex or gender identity",
      Characteristic == "Student special education status" ~ "Special education status",
      Characteristic == "Student immigration status" ~ "Immigration status",
      Characteristic == "Student english language learner status" ~ "English language learner",
      TRUE ~ Characteristic
    )
  )

## Any remaining evidence_summary rows (no indent, bolded)
rest_blk <- slice(evidence_summary, -1:-3) %>%
  transmute(
    Characteristic = str_to_sentence(Category),
    N              = N,
    Percent        = round(Percentage, 0),
    indent         = FALSE
  )

# Stack them in the correct order
evidence_tbl <- bind_rows(
  eff_block,
  acc_block,
  stake_hdr, stake_blk,
  inf_hdr,   inf_blk,
  out_hdr,   out_blk,
  eq_hdr,    eq_blk,
  rest_blk
) %>%
  mutate(
    # turn any NA into blank text
    N       = if_else(is.na(N),       "", as.character(N)),
    Percent = if_else(is.na(Percent), "", paste0(Percent, "%"))
  )

# Render gt table
evidence_tbl %>%
  gt() %>%
  cols_label(
    Characteristic = md("**Characteristic**"),
    N              = md("**N**"),
    Percent        = md("**%**")
  ) %>%
  # indent only the TRUE rows
  tab_style(
    style     = cell_text(indent = px(15)),
    locations = cells_body(rows = indent, columns = "Characteristic")
  ) %>%
  # bold all non‐indented rows
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_body(rows = !indent, columns = "Characteristic")
  ) %>%
  # one consolidated footnote on all four “*” groups
  tab_footnote(
    footnote = "* Denominator for specific categories is the number of studies on that type of evidence.",
    locations = cells_body(
      rows    = Characteristic %in% c(
        hdr_stakeholder,
        hdr_inferential,
        hdr_outcome,
        hdr_equity
      ),
      columns = "Characteristic"
    )
  ) %>%
  tab_style(
    style = cell_borders(
      sides  = "bottom",
      color  = "black",
      weight = px(2)      
    ),
    locations = cells_column_labels()
  ) %>%  
  tab_options(table.font.size = "small") %>%
  cols_hide(columns = "indent")

